---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

```

```{css}

```

## What's going on with this data?

This data set represents my observations from my daily bus commute. I take the bus for different reasons, so I decided to use that as the categorical data and the number of people I count on the bus right before I get off as the numerical data.

The reasons for my travel would be for

- Work
- University
- To meet friends
- Errands

```{r}
library(tidyverse)

#extracting data from the csv into R, then renaming the variables
logged_data <- read_csv('https://docs.google.com/spreadsheets/d/e/2PACX-1vTAF02Kzxd-nCZkoXWRuIFRNhh4TRCilwqDG5uiijEXBQpJGMS5hML1fi8MTTm5HYesBJP0kTYUbE8T/pub?output=csv') %>%
  rename(time = 1,
         count = 2,
         reason = 3)

#mean number of people on the bus over 20 days
mean_people <- ceiling(mean(logged_data$count))

#filtering for university entries
university <- logged_data %>%
  filter(reason == "University")

#filtering for everything else but uni entries
other <- logged_data %>%
  filter(reason != "University")

#mean number of people on the bus ride to university - peak travel mean
university_mean <- ceiling(mean(university$count))

#mean number of people on the bus ride during other occasions - off peak travel mean
other_mean <- ceiling(mean(other$count))


#bar chart for the total number of people who took the bus during each reason
reason_plot <- ggplot(data = logged_data) +
      geom_bar(aes(x = reason,),
               fill = "#70006c",) + 
      labs(title = "Total number of people on buses",
          subtitle = "Comparing the number of people that took the bus during different reasons of transport",
          caption = "Source: Google sheets",
          x = "Reason of travel",
          y = "Total number of people that took the bus")

#dataframe for the count of reason
reason_counts <- logged_data %>%
  count(reason)

#making a data frame that shows the frequency of each travel reason
reason_data <- data.frame(reason_counts)
colnames(reason_data) <- c("reason", "count")

#bar chart for the frequency count of each travel reason
frequency_plot <- ggplot(reason_data, aes(x = count, y = reason)) +
    geom_bar(stat = "identity", fill = "#70006c") +
    labs(title = "Frequency of Each Travel Reason",
         subtitle = "Number of times each travel reason appears",
         caption = "Source: Google Sheets",
         x = "Reason for Travel",
         y = "Frequency")

#I have decided to use mean_people, university_mean and other_mean as the summary values
#reason_plot and frequency_plot will be used as the two bar charts

```
The average number of people who take the bus overall is `r mean_people` people.
The average number of people who take the bus during peak hours and offpeak hours are `r university_mean` and `r other_mean` respectively.

### Bar chart representing the total number of people who take the bus during different travel reasons
```{r fig.cap = "Reason chart", echo = FALSE}
reason_plot
```

### Bar chart representing the different total number of times I travelled for different reasons
```{r fig.cap = "Frequency chart", echo = FALSE}
frequency_plot
```


